Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-11 troll/algs_batch.tex
small clarification to the first batch algo

Given $G(Y) =(V,E(Y))$, we have at our disposal a training set \trainset{} of labeled edges from
$E(Y)$, our goal being that of building a predictive model for the labels of the remaining edges.

@@

\begin{equation}
  \label{eq:predictor}
  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tau\Big)~,
\end{equation}

@@

\iffalse
\subsection{Approximation to Bayes via Active Learning}\label{ss:active}

@@

Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and
$\alpha \in (0,1)$ there exists a set\footnote{$E_L$ is needed to find an estimate $\tauhat$ of
$\tau$ in~(\ref{eq:predictor}) ---see Step~3 of the algorithm. Any undirected matching of $G$ of
size $\mathcal{O}(\log|V|)$ can be used, for instance using the blossom
algorithm~\autocite{matching65}. In practice, however, we never computed $E_L$, and estimated $\tau$
on the entire training set \trainset{}.} $E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex
$u \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---ie{}.,
$(u,v)$--- and at most once as destination ---ie{}., $(v,u)$. Moreover, we assume \trainset{} has
been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$. The
algorithm performs the following steps:

@@

	\item Any remaining edge $(u,v) \in E\setminus \trainset{}$ is predicted as
		$$
		\yhat(u,v) = \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) - \tfrac{1}{2} -\tauhat\Big)
		$$
\end{enumerate}

The next result\footnote{All proofs are in the supplementary material.} shows that if the graph is
not too sparse, then the above algorithm can approximate the Bayes optimal predictor on nodes whose
in-degree and out-degree is not too small.

\begin{theorem}
  \label{t:active}
  Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the
  model in \autoref{s:gen}.
  If the algorithm is run with parameter $Q = \Omega(\ln|V|)$, and $\alpha \in (0,1)$ such that the
  above assumptions are satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
  simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\ln|V|)$, and
  $\eta(u,v) = \Pr(\yuv=1)$ is bounded away from $\tfrac{1}{2}$.
  The overall number of training edges is $\Theta\big(|V|\ln|V|\big)$.
\end{theorem}

The approach leading to Theorem~\ref{t:active} lets us derive the \usrule{} algorithm assessed in
our experiments of \autoref{s:exp}, but it needs the graph to be sufficiently dense and the bias
$\tau$ to be the same for all edges. In order to address these limitations, we now introduce a
second method based on label propagation.

@@

Binary $\pm 1$ predictions on the test set $E\setminus \trainset{}$ are then obtained by
thresholding the obtained values $\yuv$ at $0$.

>>>>

Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
labeled edges from $E(Y)$. We want to build a predictive model for the labels of the remaining
edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
locally the parameters by their empirical means in the training set, which under some density
assumptions are showed to concentrate around their true values. The second algorithm, on the other
hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
parameters through a global label propagation approach, without making any density assumption.

\subsection{Approximation to Bayes via dense sampling}\label{ss:bayes_approx}

@@

\begin{equation}
  \label{eq:predictor}
  \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)~,
\end{equation}

@@

\iffalse

@@

Our algorithm works under the assumption that for given parameters $Q$ (a positive integer) and
$\alpha \in (0,1)$, there exists a set\footnote{$E_L$ is needed to find an estimate $\tauhat$ of
$\tau$ in~(\ref{eq:predictor}), as done in Step~3 of the algorithm. Any undirected matching of $G$ of
size $O(\log|V|)$ can be used, obtained for instance by the blossom algorithm~\autocite{matching65}.
In practice, however, we never computed $E_L$, and estimated $\tau$
on the entire training set \trainset{}.} $E_L \ss E$ of size $\tfrac{2Q}{\alpha}$ where each vertex
$u \in V$ appearing as an endpoint of some edge in $E_L$ occurs at most once as origin ---\ie{}
$(u,v)$--- and at most once as destination ---\ie{} $(v,u)$. Moreover, we assume \trainset{} has
been drawn from $E$ at random {\em without} replacement, with $m = |E_0| = \alpha\,|E|$. The
algorithm performs the following steps:

@@

	\item Any remaining edge $(u,v) \in E\setminus \trainset{}$ is predicted as
		$$
		\yhat(u,v) = \sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tauhat - \tfrac{1}{2}\Big)
		$$
\end{enumerate}

The next result shows that the above algorithm can approximate the Bayes optimal predictor on nodes
whose in-degree and out-degree is not too small.

\begin{theorem}
  \label{t:active}
  Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the
  model in \autoref{s:gen}.
  If the algorithm is run with parameter $Q = \Omega(\log|V|)$, and $\alpha \in (0,1)$ such that the
  above assumptions are satisfied, then $\yhat(u,v) = y^*(u,v)$ holds with high probability
  simultaneously for all test edges $(u,v) \in E$ such that $\dout(u),\din(v) = \Omega(\log|V|)$, and
  $\eta(u,v) = \Pr(\yuv=1)$ is bounded away from $\tfrac{1}{2}$.
\end{theorem}

While we defer the full proof of \autoref{t:active} to the additional material at the end of this
chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used, focusing on
parameter $p$. First, by \autoref{l:bins} on negatively associated random variables (stated
\vpageref{l:bins}), we show that with our choice of $Q$, for $\theta=\frac{2Q}{\alpha}$ and for any
$u$ having enough out neighbors (\ie{} $\NNout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
are in the training set with high probability.
We use these $Q$ edges to compute $1-\htr(u)$, the empirical probability of drawing a $+1$-labeled
edge from $\Nout(i)$. We show that it is close to the real probability, which according to equation
\eqref{e:pout} is equal to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$. $\qbar_u$ being a sample mean
of i.i.d.~$[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1
\mu\big(p,\cdot\big) dp$, it concentrates around its expectation $\mu_q$.

By following the same reasoning on the parameters $q$ and using an union bound, we then show that
$\left(1-\htr(u)\right) + \left(1-\hun(v)\right) - \frac{\mu_p+\mu_q}{2}$ is close to
$\frac{p_u+q_v}{2}$ simultaneously for all edges. Another set of concentration arguments then show
that our estimate $\tauhat$ is close to $\frac{\mu_p+\mu_q}{2}$. Combining these two facts ends the
proof.

\medskip

The approach leading to \autoref{t:active} needs the graph to be sufficiently dense. At first sight,
having $Q = \Omega(\log|V|)$ training edges per nodes does not sound that like such a stringent
requirement. Consider for instance Facebook ---which is neither signed nor directed though. It has
two billion
users~\bottomfootnote{\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
each of them having $155\approx 7.2\log|V|$ friends on average~\autocite{facebookFriend16}. However,
the constant hidden in the $\Omega$ notation, and needed to guarantee a good behavior, is much
larger than $7.2$. We will nonetheless see in the experiments that we can still apply the resulting
algorithm (we call it \usrule{}) with satisfying results. Additionally, and in order to address this
density limitation, we now introduce a second method based on label propagation.

@@

Binary $\pm 1$ predictions on the test set $E\setminus \trainset{}$ are then obtained by
thresholding the computed values $\yuv$ at $0$.

### 2017-09-11 troll/experiments.tex
small clarification to the first batch algo

  \item The algorithm analyzed at the beginning of \autoref{s:algbatch}, which we call \usrule{}
    (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
    computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all $u \in V$ (or setting those
    values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use
    equation \eqref{eq:predictor} and estimate $\tau$ on $E_0$.

@@

  \item As claimed in the introduction, our Bayes approximator \usrule{} closely mirrors in
    performance the more involved \uslogregp{} model. In fact, supporting our generative model of
    \autoref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$
    are almost equal (see Table~2 in the supplementary material), thereby suggesting that
    predictor~\eqref{eq:predictor}, derived from the theoretical results at the beginning of
    \autoref{s:algbatch}, is \emph{also} the best logistic model based on trollness and
    untrustworthiness.
\end{enumerate}

>>>>

  \item The algorithm analyzed \autoref{ss:bayes_approx}, which we call \usrule{}
    (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
    computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all $u \in V$ (or setting those
    values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use
    equation \eqref{eq:predictor} and estimate $\tau$ on $E_0$.

@@

  \item As claimed in the introduction, our Bayes approximator \usrule{} closely mirrors in
    performance the more involved \uslogregp{} model. In fact, supporting our generative model of
    \autoref{s:gen}, the logistic regression weights for features $1-\htr(i)$ and $1-\hun(j)$
    are almost equal (see Table~2 in the supplementary material), thereby suggesting that
    predictor~\eqref{eq:predictor}, derived from the theoretical results in
    \autoref{ss:bayes_approx}, is \emph{also} the best logistic model based on trollness and
    untrustworthiness.
\end{enumerate}

### 2017-09-11 troll/generative.tex
small clarification to the first batch algo

Notice that the Bayes optimal prediction for $\yuv$ is $y^*(u,v) = \sgn\big(\eta(u,v) -
\tfrac{1}{2}\big)$, where $\eta(u,v) = \Pr\big( \yuv = 1 \big)$. Moreover, the probability of
drawing at random a $+1$-labeled edge from $\Nout(u)$ equals
\begin{equation}\label{e:pout}
	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \NNout(u)} q_v \Biggl)
\end{equation}
Similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(v)$ equals
\begin{equation}\label{e:pin}
	\frac{1}{2}\,\Biggl(q_v + \frac{1}{\din(v)} \sum_{u\in \NNin(v)} p_u \Biggl)
\end{equation}

>>>>

Notice that the Bayes optimal prediction for $\yuv$ is $y^*(u,v) = \sgn\big(\eta(u,v) -
\tfrac{1}{2}\big)$, where $\eta(u,v) = \Pr\big( \yuv = 1 \big)$. Moreover, the probability of
drawing at random a $+1$-labeled edge from $\Nout(u)$ equals
\begin{equation}\label{e:pout}
	\frac{1}{2}\,\Biggl(p_u + \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v \Biggl) =
		\frac{1}{2}\,\left(p_u + \qbar_u\right)\,,
\end{equation}
where $\qbar_u = \frac{1}{\dout(u)} \sum_{v\in \nnout(u)} q_v$ is the average $q$ value of $u$ out
neighbors.
Similarly, the probability of drawing at random a $+1$-labeled edge from $\Nin(v)$ equals
\begin{equation}\label{e:pin}
	\frac{1}{2}\,\Biggl(q_v + \frac{1}{\din(v)} \sum_{u\in \NNin(v)} p_u \Biggl)=
		\frac{1}{2}\,\left(q_v + \pbar_v\right)
\end{equation}
